---
title: "Power simulations"
author: "Benjamin Chan (chanb@ohsu.edu)"
date: `r Sys.Date()`
output: html_document
---

```{r setup, include=FALSE}
setwd("C:/Users/chanb/Box Sync/Share/BDP2-2 Horner-Johnson/scripts")
library(checkpoint)
library(knitr)
library(rmarkdown)
checkpoint("2018-04-01", use.knitr = TRUE)
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      fig.path = "../figures/", 
                      dpi = 300)
library(magrittr)
library(dplyr)
library(broom)
library(parallel)
library(doParallel)
library(ggplot2)
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
Sys.time0 <- Sys.time()
```

## Generate data

```{r}
N <- 600
S <- 100
```

Study design is a two-factor, full factorial randomization.
The factors are **framing** and the **message** of the intervention.
Each factor has **3** levels.

Full sample size is **`r sprintf("N = %d", N)`**.

Number of simulations is **`r sprintf("S = %d", S)`**.

The outcomes are **comprehension** and **influence**.

```{r}
comprehension0 <- 0.5
influence0 <- 0
sd0 <- 1
```

Comprehension is a dichotomous outcome.
Baseline comprehension is assumed to be **`r sprintf("%.00f%%", comprehension0 * 100)`**.

Influence is a continuous outcome.
It represents pre-post change on a scale.
For analytic purposes, it will be scaled on a (-1, +1) continuum.
Baseline influence is assumed to be **`r sprintf("%.02f", influence0)`**.

```{r generateData}
df <- 
  expand.grid(framing = c("A", "B", "C"),
              message = c("X", "Y", "Z"),
              id = seq(round(N / 9)),
              sample = seq(S)) %>% 
  mutate(eta1 = 
           log(comprehension0 / (1 - comprehension0)) + 
           as.integer(framing == "A") *  0.0 + 
           as.integer(framing == "B") *  0.5 + 
           as.integer(framing == "C") *  0.67 + 
           as.integer(message == "X") *  0.0 + 
           as.integer(message == "Y") *  0.6 + 
           as.integer(message == "Z") *  0.75) %>% 
  mutate(comprehension = rbinom(nrow(.), 1, exp(eta1) / (1 + exp(eta1)))) %>% 
  mutate(eta2 = 
           influence0 + 
           as.integer(framing == "A") * 0.0 + 
           as.integer(framing == "B") * 0.1 + 
           as.integer(framing == "C") * 0.3 + 
           as.integer(message == "X") * 0.0 + 
           as.integer(message == "Y") * 0.2 + 
           as.integer(message == "Z") * 0.4) %>% 
  mutate(influence = rnorm(nrow(.), eta2, sd = sd0))
dfSum <- 
  df %>% 
  group_by(framing, message, sample) %>% 
  summarize(n = n(),
            eta1 = mean(eta1),
            y1 = sum(comprehension),
            p1 = sum(comprehension) / n(),
            eta2 = mean(eta2),
            y2 = mean(influence),
            s2 = sd(influence))
dfSum %>% 
  group_by(framing, message) %>% 
  summarize(n = n(),
            meanComprehension = mean(p1),
            nominalOddsRatio = exp(mean(eta1)),
            nominalRelRisk = exp(mean(eta1)) / (1 - comprehension0 + (comprehension0 * exp(mean(eta1)))),
            meanInfluence = mean(y2),
            meanSD = mean(s2),
            nominalEffectSize = mean(eta2)) %>% 
  select(-nominalOddsRatio) %>% 
  kable(digits = 3)
dfSum %>% 
  ggplot(aes(x = p1)) +
  geom_density(fill = "grey", alpha = 1/2) +
  scale_x_continuous("Comprehension") +
  facet_grid(message ~ framing)
dfSum %>% 
  ggplot(aes(x = y2)) +
  geom_density(fill = "grey", alpha = 1/2) +
  scale_x_continuous("Mean influence") +
  facet_grid(message ~ framing)
dfSum %>% 
  ggplot(aes(x = s2)) +
  geom_density(fill = "grey", alpha = 1/2) +
  scale_x_continuous("SD influence") +
  facet_grid(message ~ framing)
```

## Aim 1: Comprehension

```{r comprehension}
cores <- min(4, detectCores() - 1)
cl <- makeCluster(cores)
registerDoParallel(cl)
simBetas <- foreach (i = seq(S), .combine = rbind) %dopar% {
  require(magrittr)
  require(dplyr)
  require(broom)
  df %>% 
    filter(sample == i) %>% 
    glm(comprehension ~ framing + message, data = ., family = "binomial") %>% 
    tidy() %>% 
    mutate(sim = i)
}
stopCluster(cl)
power <- 
  simBetas %>% 
  filter(term != "(Intercept)") %>% 
  mutate(sig = p.value < 0.05)
power %>% 
  group_by(term) %>% 
  summarize(n = n(),
            truePos = sum(sig)) %>% 
  mutate(power = truePos / n) %>% 
  kable()
power %>% 
  ggplot(aes(x = estimate, fill = sig)) +
  geom_histogram(alpha = 1/2) +
  scale_fill_discrete("", labels = c("False Neg", "True Pos")) +
  facet_wrap(~ term)
```


## Aim 2: Influence

```{r}
cores <- min(4, detectCores() - 1)
cl <- makeCluster(cores)
registerDoParallel(cl)
simBetas <- foreach (i = seq(S), .combine = rbind) %dopar% {
  require(magrittr)
  require(dplyr)
  require(broom)
df %>% 
    filter(sample == i) %>% 
  lm(influence ~ framing + message, data = .) %>% 
    tidy() %>% 
    mutate(sim = i)
}
stopCluster(cl)
power <- 
  simBetas %>% 
  filter(term != "(Intercept)") %>% 
  mutate(sig = p.value < 0.05)
power %>% 
  group_by(term) %>% 
  summarize(n = n(),
            truePos = sum(sig)) %>% 
  mutate(power = truePos / n) %>% 
  kable()
power %>% 
  ggplot(aes(x = estimate, fill = sig)) +
  geom_histogram(alpha = 1/2) +
  scale_fill_discrete("", labels = c("False Neg", "True Pos")) +
  facet_wrap(~ term)
```


## Session summary

```{r}
list(completionDateTime = Sys.time(),
     executionTime = Sys.time() - Sys.time0,
     sessionInfo = sessionInfo())
```
